---
title: "Part II Conditional Hypothesis"
author: "Sindhi Polubothu"
---

```{r setup, include=FALSE}
setwd("C:/Users/Spolubot/Documents/Classes/Data Science/Assignment 2")

data <- read.csv("AllViolenceData_170216.csv")
```

##Part II Conditional Hypothesis

__Hypothesis__:  
Our conditional hypothesis is that the total number of people dead is conditional upon the interaction between which federal force was involved and how many people were detained. We predict that if the army/navy/federal police is involved in an event each additional person detained results in a lower number of total people dead on average. The rational behind this prediction is that if there an event with federal forces ends with detained people, it is likely to be less deadly. 

Below is the multi-variate regression model with interactions to test this hypothesis: 

__Assumptions:__  
__navy__, __army__, and __federal police__ are binary variables in which 1 = they were involved in an event and 0 = they were not involved in an event. __total.people.dead__ and __detained__ are numeric variables in that increase by one from 0 up each representing an additonal person dead or detained. Additionally in order to preform the analysis we assumed that in each event only the army, navy, or federal police were present alone at each event. 


```{r  echo = FALSE}
lm1 <- lm(total.people.dead ~ navy + army + federal.police + detained + detained * navy + detained * army + detained * federal.police, data = data)

summary(lm1)

```

####__Marginal Effects__

__Intercept__ 
The intercept in the regression output above shows that if no navy, army, or federal police were involved an if no one was detained, the average number of total people dead is 1.03. 

__Coeffiencts__ 
Given this is a conditional interaction, the model shows that if the navy is involved and no one is detained on average .43 (95% confidence interval range of .06 to .84) more people  killed overall, holding all other variables constant with statistical signifiance (p-value of at least .05). If the army is involved and no one is detained on average .37 (95% confidence interval range of .25 to .52) more people  killed overall, holding all other variables constant with statistical signifiance (p-value of at least .05). If the federal police is involved and no one is detained on average -.25 (95% confidence interval range of -.47 to -.05) less people killed overall, holding all other variables constant with statistical signifiance (p-value of at least .05). Additionally for each additional detained person when no one from the navy, army, or federal police is involved on average .04 (95% confidence interval range of .02 to .06) more people killed overall, holding all other variables constant with statistical signifiance (p-value of at least .05).

__Interactions__ 
The interation terms show that if the navy is involved, each additional detained person recorded in the event results in on average .17 (95% confidence interval of .07 to .26) more people killed overall , holding all other variables constant with statistical signifiance (p-value of at least .05). If the army is involved, each additional detained person recorded in the event results in on average -.09 (95% confidence interval of -.15 to -.03) less people killed overall , holding all other variables constant with statistical signifiance (p-value of at least .05). Finally if the federal police are involved each additional detained person recorded in the event results in on average .06 (95% confidence interval of .00 to .12) more people killed overall, holding all other variables constant with statistical signifiance (p-value of at least .05).  

__Conclusion__ 
The hypothesis stated proved to be partial true per the data. In the case of the navy and federal police, each indivial detained person recorded actual resulted in an increase in average overall deaths which is contrary to the prediction. In the case of the army, each individual detained person recorded resulted in a decrease in average overall deaths. 

__Computations__ 
The marginal effect referenced was the "Estimate" column provided by the regression model and is interpreted as described in the marginal effect section above. To arrive at standard error and confidence intervals I used the "Standard Error" Column provided by regression and multipied by 2 and -2 and added to the "Estimate" provided to arrive at a 95% confidence interval. 

__Limitations__ 
The data is limited as the data is only provides information for a few year period of time in certain locations in mexico. The regression focuses on total people dead and did not segregate the analysis by which type of person died in each event.As such more detailed analysis would have to be preformed to get more insight on what type of person dies as more people are detained by certain armed forces. Additionally it is possible that the army, navy, and federal police were all at certain events which would complicate the analysis of the data as they are interrelated. 

__Code__ 

```{r}
setwd("C:/Users/Spolubot/Documents/Classes/Data Science/Assignment 2")

data <- read.csv("AllViolenceData_170216.csv")

lm1 <- lm(total.people.dead ~ navy + army + federal.police + detained + detained * navy + detained * army + detained * federal.police, data = data)

summary(lm1)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
